Transient Corotating Clumps Around Adolescent Low-Mass Stars: Four Years of Complex Quasiperiodic Variables from TESS

### 2023-08-17 paper/ms.tex
update ms

Two exceptions to the rule of ``simple field topologies for magnetic B
stars'' are HD~37776 and HD~64740, which are B stars that were
previously known from spectropolarimetry to have field geometries
dominated by high order multipoles \citep{2011ApJ...726...24K}.
Reviewing the literature, we learned that recent TESS light curves of
these two B stars appear surprisingly similar to the CQV light curves
\citep{2020pase.conf...46M}.  The middle row of Figure~\ref{fig:bstar}
shows the phased TESS light curves for these two outlying magnetic B
stars, with by-eye ``best matching'' CQVs shown for comparison.  The
number of dips per cycle, the shapes of the dips, and the dip depths
relative to the sinusoidal envelope all appear to be similar.  The
implication of this surprising connection is that the
highly-structured M dwarf light curves may imply multipolar magnetic
fields -- since it is the complex fields which are the defining
characteristic of these two B stars, relative to other known magnetic
B stars like $\sigma$~Ori~E.  This is consistent with the
non-axisymmetric topologies that have been reported in a small number
of M dwarfs for which Zeeman Doppler Imaging is technically feasible
\citep[see the review by][]{2021A&ARv..29....1K}.

The physical similarity between the B stars and the M dwarfs
presumably has its origin in the existence of a ``centrifugal
magnetosphere'' \citep[see][]{2013MNRAS.429..398P}.  In other words,
both classes of objects probably satisfy the condition $R_{\rm A} >
R_{\rm cr}$, for $R_{\rm A}$ the Alfv\'en radius.  When this condition
holds, stable points exist wherever the outward centrifugal force is
balanced by the combined effect of gravity and magnetic tension along
closed field loops \citep{1985Ap&SS.116..285N}.  In the converse
``dynamical'' case, when  $R_{\rm A} < R_{\rm cr}$, any material
launched by the star returns to its surface over the free-fall
timescale.  A simple order of magnitude estimate assuming a dipole
field with $B_0\approx1\,{\rm kG}$ at the star's surface, a local
plasma number density $n\approx10^9\,{\rm cm^{-3}}$, and a plasma
temperature $10^6\,{\rm K}$ gives Alfv\'en radii of order a few times
the corotation radii, $R_{\rm cr}$.  This suggests that the existence
of a centrifugal magnetosphere is plausible in young, rapidly rotating
M dwarfs.

@@

	\vspace{-1.1cm}
	\subfloat{
			\includegraphics[width=0.9\textwidth]{f13c.pdf}
    }
	\end{center}
	\vspace{-0.4cm}
	\caption{
		{\bf Spectral age and activity diagnostics for three CQVs}.
		Wavelengths are in air;
		the continuum normalization is relative to the entire order.
		The H$\alpha$ emission strength classifies the stars as weak-lined
		T Tauris.
		The lithium detection for TIC~146539195 is consistent with its
		mass and $\beta$~Pic membership;
		the non-detections for TIC~402980664 and TIC~264599508
		are consistent with the $\approx$42\,Myr age implied by their membership
		in the Columba moving group.
	}
	\label{fig:tic3006timegroups}
\end{figure*}

>>>>

Two exceptions to the rule of ``simple field topologies for magnetic B
stars'' are HD~37776 and HD~64740, which are B stars that were
previously known from spectropolarimetry to have field geometries
dominated by high order multipoles \citep{2011ApJ...726...24K}.
Reviewing the literature, we learned that recent TESS light curves of
these two B stars appear surprisingly similar to the CQV light curves
\citep{2020pase.conf...46M}.  The middle row of Figure~\ref{fig:bstar}
shows the phased TESS light curves for these two outlying magnetic B
stars, with by-eye ``best matching'' CQVs shown underneath for comparison.  The
number of dips per cycle, the shapes of the dips, and the dip depths
relative to the sinusoidal envelope all appear to be similar.  This surprising connection 
suggests that the
highly-structured M dwarf light curves may imply multipolar magnetic
fields -- since it is the complex fields which are the defining
characteristic of these two B stars, relative to other known magnetic
B stars like $\sigma$~Ori~E.  This is consistent with the
non-axisymmetric topologies that have been reported in a small number
of M dwarfs for which Zeeman Doppler Imaging is technically feasible
\citep[see][and references therein]{2021A&ARv..29....1K}.

The physical similarity between the B stars and the M dwarfs
presumably has its origin in the existence of a ``centrifugal
magnetosphere'' \citep[see][]{2013MNRAS.429..398P}.  In other words,
both classes of objects probably satisfy the condition $R_{\rm A} >
R_{\rm cr}$, for $R_{\rm A}$ the Alfv\'en radius.  When this condition
holds, stable points exist wherever the outward centrifugal force is
balanced by the combined effect of gravity and magnetic tension along
closed field loops \citep{1985Ap&SS.116..285N}.  In the converse
``dynamical'' case, when  $R_{\rm A} < R_{\rm cr}$, any material
launched by the star returns to its surface over the free-fall
timescale.  A simple estimate assuming a dipole
field with $B_0\approx1\,{\rm kG}$ at the star's surface, a local
plasma number density $n\approx10^9\,{\rm cm^{-3}}$, and a plasma
temperature $10^6\,{\rm K}$ gives Alfv\'en radii of order a few times
the corotation radii, $R_{\rm cr}$.  This suggests that the existence
of a centrifugal magnetosphere is plausible in young, rapidly rotating
M dwarfs.

@@

	\vspace{-1.1cm}
	\subfloat{
			\includegraphics[width=0.9\textwidth]{f13c.pdf}
    }
	\end{center}
	\vspace{-0.4cm}
	\caption{
		{\bf Spectral age and activity diagnostics for three CQVs}.
		Wavelengths are in air;
		the continuum normalization is relative to the entire order.
		The H$\alpha$ emission strength classifies the stars as weak-lined
		T Tauris.
		The lithium detection for TIC~146539195 is consistent with its
		mass and $\beta$~Pic membership;
		the non-detections for TIC~402980664 and TIC~264599508
		are consistent with the $\approx$42\,Myr age implied by their membership
		in the Columba moving group.
	}
	\label{fig:speccutouts}
\end{figure*}
